---
title: 'Intermediate Deliverable: Plots'
author: "Angel Sarmiento"
date: "3/18/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, comments = NA, message = FALSE)
```


```{r, echo = FALSE, warning = FALSE, comments = NA}
#library Import
library(caret)
library(tidyverse)

library(fable)
library(feasts)
library(fredr)
library(tsibble)
library(patchwork)
library(kableExtra)
library(ggfortify)

set.seed(23)
data <-  read_csv("dataset.csv")  %>% na.pass()

colnames(data)[2:7] <- c("nonfarm", "eci", "avg_week_hrs", "avg_hr_earnings",
                         "avg_week_earnings", "all_goods")
data["tot_week_earnings"] <- data[5]*data[2]
```

# Introduction

There are a number of MSA's in Florida that have a pretty significant number of metrics used to describe and analyse them in a macroeconomical sense. MSA's are Metropolitan Statistical Areas, a designation given for areas with largely dense populations that have close economic ties to their surrounding areas. For this investigation, the Miami-West Palm-Fort Lauderdale MSA will be analysed with the final intent of delivering the best model possible for predicting the total nonfarm employment. The different models illustrated in this investigation all involve adjustments and reflections of the time series quality of the data. The best model will be chosen from a host of other generated models that will be compared with a number of comparison statistics. Namely: AIC, BIC, RMSE, and Out-of-sample RMSE. 
   
\qquad The relevant predictors include the Economic Conditions Index (ECI): A value that shows the economic growth of an MSA, the average weekly hours worked, the average hourly earnings of employees, the average weekly earnings, all of the good produced, and the total weekly earnings of all employees. The economic conditions index was chosen because it makes intutive sense that the growth of an economy could contribute to employment in that economy. It also makes intuitive sense that there could be some positive relationship between the nonfarm employment and the amount of goods produced. 


# Analysis of Variables

```{r, warning = FALSE, comments = NA, message = FALSE}
summary(data[2:5]) %>%
  kable(format = "latex") %>% 
  row_spec(0,bold=TRUE) %>% 
  kable_styling(full_width = FALSE, position = "center", latex_options = c("striped", "HOLD_position"))

summary(data[6:7]) %>%
  kable(format = "latex") %>% 
  row_spec(0,bold=TRUE) %>% 
  kable_styling(full_width = FALSE, position = "center", latex_options = c("striped", "HOLD_position"))
```

\begin{center}
Table 1. Summary statistics of all of the variables in the analysis
\end{center}
```{r}
dataset <- data %>% as_tsibble()
#creating the months variable
dataset <- dataset %>%
  mutate(Month = lubridate::month(DATE)) %>%
  index_by(Month) 
```

```{r, fig.height = 8, fig.width = 9}
t1 <- dataset %>% autoplot(nonfarm, ts.colour = "red") +
  xlab("Date") +
  ylab("Nonfarm")
t2 <- dataset %>% autoplot(eci) +
  xlab("Date") +
  ylab("ECI")
t3 <- dataset %>% autoplot(avg_week_hrs) +
  xlab("Date") +
  ylab("Avg Weekly Hours")
t4 <- dataset %>% autoplot(avg_hr_earnings) +
  xlab("Date") +
  ylab("Avg Hourly Earnings")
t5 <- dataset %>% autoplot(avg_week_earnings)+
  xlab("Date") +
  ylab("Avg Weekly Earnings")
t6 <- dataset %>% autoplot(all_goods)+
  xlab("Date") +
  ylab("All Goods Produced")
t7 <- dataset %>% autoplot(tot_week_earnings)+
  xlab("Date") +
  ylab("Total Weekly Earnings")

(t1 + t2) /
(t3 + t4) /
(t5 + t6) /
(t7)
```
\begin{center}
Figure 1. Time Series plots of every variable.
\end{center}


```{r, fig.height = 6, fig.width = 10}

ac_data <- dataset %>% na.omit()

a1 <- autoplot(acf(ac_data$nonfarm, plot = FALSE), main = "Nonfarm") 
a2 <- autoplot(acf(ac_data$eci, plot = FALSE), main = "ECI") + theme(axis.title.y = element_blank())
a3 <- autoplot(acf(ac_data$avg_week_hrs, plot = FALSE), main = "Weekly Hours")  + theme(axis.title.y = element_blank())
a4 <- autoplot(acf(ac_data$avg_hr_earnings, plot = FALSE), main = "Hourly Earnings")
a5 <- autoplot(acf(ac_data$avg_week_earnings, plot = FALSE), main = "Weekly Earnings")  + theme(axis.title.y = element_blank())
a6 <- autoplot(acf(ac_data$all_goods, plot = FALSE), main = "All Goods Produced ")  + theme(axis.title.y = element_blank())
a7 <- autoplot(acf(ac_data$tot_week_earnings, plot = FALSE), main = "Total Weekly Earnings")


wrap_plots(a1, a2, a3 , a4, a5, a6, a7, ncol = 3, byrow = TRUE)
```
\begin{center}
Figure 2. Autocorrelograms of all variables 
\end{center}

```{r, fig.height = 7, fig.width = 10}
p1 <- autoplot(pacf(ac_data$nonfarm, plot = FALSE), main = "Nonfarm") 
p2 <- autoplot(pacf(ac_data$eci, plot = FALSE), main = "ECI") + theme(axis.title.y = element_blank())
p3 <- autoplot(pacf(ac_data$avg_week_hrs, plot = FALSE), main = "Weekly Hours")  + theme(axis.title.y = element_blank())
p4 <- autoplot(pacf(ac_data$avg_hr_earnings, plot = FALSE), main = "Hourly Earnings")
p5 <- autoplot(pacf(ac_data$avg_week_earnings, plot = FALSE), main = "Weekly Earnings")  + theme(axis.title.y = element_blank())
p6 <- autoplot(pacf(ac_data$all_goods, plot = FALSE), main = "All Goods Produced ")  + theme(axis.title.y = element_blank())
p7 <- autoplot(pacf(ac_data$tot_week_earnings, plot = FALSE), main = "Total Weekly Earnings")


wrap_plots(p1, p2, p3 , p4, p5, p6, p7, ncol = 3, byrow = TRUE)
```
\begin{center}
Figure 3. Partial Autocorrelograms of all variables
\end{center}


\cleardoublepage
# Appendix A: Code

```{r, eval = FALSE, echo = TRUE}
## ----setup, include=FALSE-----------------------------------------------------------------------------------------------------------------
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, comments = NA, message = FALSE)


## ---- echo = FALSE, warning = FALSE, comments = NA----------------------------------------------------------------------------------------
#library Import
library(caret)
library(tidyverse)

library(fable)
library(feasts)
library(fredr)
library(tsibble)
library(patchwork)
library(kableExtra)
library(ggfortify)

set.seed(23)
data <-  read_csv("dataset.csv")  %>% na.pass()

colnames(data)[2:7] <- c("nonfarm", "eci", "avg_week_hrs", "avg_hr_earnings",
                         "avg_week_earnings", "all_goods")
data["tot_week_earnings"] <- data[5]*data[2]


## ---- warning = FALSE, comments = NA, message = FALSE-------------------------------------------------------------------------------------
summary(data[2:5]) %>%
  kable(format = "latex") %>% 
  row_spec(0,bold=TRUE) %>% 
  kable_styling(full_width = FALSE, position = "center", latex_options = c("striped", "HOLD_position"))

summary(data[6:7]) %>%
  kable(format = "latex") %>% 
  row_spec(0,bold=TRUE) %>% 
  kable_styling(full_width = FALSE, position = "center", latex_options = c("striped", "HOLD_position"))


## -----------------------------------------------------------------------------------------------------------------------------------------
dataset <- data %>% as_tsibble()
#creating the months variable
dataset <- dataset %>%
  mutate(Month = lubridate::month(DATE)) %>%
  index_by(Month) 


## ---- fig.height = 8, fig.width = 9-------------------------------------------------------------------------------------------------------
t1 <- dataset %>% autoplot(nonfarm, ts.colour = "red") +
  xlab("Date") +
  ylab("Nonfarm")
t2 <- dataset %>% autoplot(eci) +
  xlab("Date") +
  ylab("ECI")
t3 <- dataset %>% autoplot(avg_week_hrs) +
  xlab("Date") +
  ylab("Avg Weekly Hours")
t4 <- dataset %>% autoplot(avg_hr_earnings) +
  xlab("Date") +
  ylab("Avg Hourly Earnings")
t5 <- dataset %>% autoplot(avg_week_earnings)+
  xlab("Date") +
  ylab("Avg Weekly Earnings")
t6 <- dataset %>% autoplot(all_goods)+
  xlab("Date") +
  ylab("All Goods Produced")
t7 <- dataset %>% autoplot(tot_week_earnings)+
  xlab("Date") +
  ylab("Total Weekly Earnings")

(t1 + t2) /
(t3 + t4) /
(t5 + t6) /
(t7)


## ---- fig.height = 6, fig.width = 10------------------------------------------------------------------------------------------------------

ac_data <- dataset %>% na.omit()

a1 <- autoplot(acf(ac_data$nonfarm, plot = FALSE), main = "Nonfarm") 
a2 <- autoplot(acf(ac_data$eci, plot = FALSE), main = "ECI") + theme(axis.title.y = element_blank())
a3 <- autoplot(acf(ac_data$avg_week_hrs, plot = FALSE), main = "Weekly Hours")  + theme(axis.title.y = element_blank())
a4 <- autoplot(acf(ac_data$avg_hr_earnings, plot = FALSE), main = "Hourly Earnings")
a5 <- autoplot(acf(ac_data$avg_week_earnings, plot = FALSE), main = "Weekly Earnings")  + theme(axis.title.y = element_blank())
a6 <- autoplot(acf(ac_data$all_goods, plot = FALSE), main = "All Goods Produced ")  + theme(axis.title.y = element_blank())
a7 <- autoplot(acf(ac_data$tot_week_earnings, plot = FALSE), main = "Total Weekly Earnings")


wrap_plots(a1, a2, a3 , a4, a5, a6, a7, ncol = 3, byrow = TRUE)


## ---- fig.height = 7, fig.width = 10------------------------------------------------------------------------------------------------------
p1 <- autoplot(pacf(ac_data$nonfarm, plot = FALSE), main = "Nonfarm") 
p2 <- autoplot(pacf(ac_data$eci, plot = FALSE), main = "ECI") + theme(axis.title.y = element_blank())
p3 <- autoplot(pacf(ac_data$avg_week_hrs, plot = FALSE), main = "Weekly Hours")  + theme(axis.title.y = element_blank())
p4 <- autoplot(pacf(ac_data$avg_hr_earnings, plot = FALSE), main = "Hourly Earnings")
p5 <- autoplot(pacf(ac_data$avg_week_earnings, plot = FALSE), main = "Weekly Earnings")  + theme(axis.title.y = element_blank())
p6 <- autoplot(pacf(ac_data$all_goods, plot = FALSE), main = "All Goods Produced ")  + theme(axis.title.y = element_blank())
p7 <- autoplot(pacf(ac_data$tot_week_earnings, plot = FALSE), main = "Total Weekly Earnings")


wrap_plots(p1, p2, p3 , p4, p5, p6, p7, ncol = 3, byrow = TRUE)
```



